Analysing Melbourne CBD pedestrian counts data with a focus on imputation methods

### 2017-10-21 thesis_draft/01-chap1.Rmd
bookdown update - need to fix formatting, improve introduction, write abstract, finish off lit review and conclusion

---
title: "Analysing Melbourne CBD pedestrian counts data with a focus on imputation methods"
author: "Gavin Chin"
date: "21 September 2017"
output:
  pdf_document: default
  html_document: default
  word_document: default
fontsize: 12pt
---

@@

dfa$DayType <-       ifelse(dfa$Day == "Sunday", "Sunday",
                     ifelse(dfa$Day == "Saturday", "Saturday",
                    ifelse(dfa$Day == "Monday", "Monday",
                     ifelse(dfa$Day == "Friday", "Friday",
                                  "Midweek"))))
dfa$DayType[dfa$IsHDay == 1] <- "Holiday"
dfa$HDay <- ifelse(dfa$IsHDay == 1, "Holiday", dfa$Day)
original_df <- dfa
dfa <- filter(dfa, Date_Time < "2017-01-01")
ori_2 <- filter(original_df, Date_Time < "2017-01-01")
ori_ts <- filter(original_df, Date_Time > "2016-12-31")
ori_ts2 <- ori_ts
```
## Background

#### The Data: City of Melbourne Pedestrian Data

@@

### Literature Review
<!-- #### IMPUTATION OF MISSING CLASSIFIED TRAFFIC DATA DURING WINTER SEASON - _Hyuk-Jae Roh, Satish Sharma, Prasanta K. Sahu_   -->

@@

## Methodology: Imputation of Missing Values

@@

```{r naprop_hist, fig.width=9, fig.height=4, fig.cap="Histograms of the proportions of missing values at each sensor during 2014 - 2016 when calculated before and after correction for false zero counts. The distributions of proportion of missing values at each sensor is significantly different after correcting for false zero counts in the data. This emphasises the necessity to correct for false zero counts as it will affect "}
theme_g <- theme_minimal() +
           theme(plot.title = element_text(hjust = 0.5,
                                           size = 10))
p1 <- ggplot() + geom_histogram(aes(x = na_prop*100), bins = 12) +
          xlab("Percentage of Missing Values (%)") +
          ylab("Number of Sensors") + theme_g +
          geom_vline(xintercept = 10, size = 2,
                     colour = 'red') +
          geom_label(aes(label = "10% Threshold", x = 16, y = 25),
                    colour = "white", fill = 'black', alpha = 0.4) +
          scale_y_continuous(limits = c(0,35))
p2 <- ggplot() + geom_histogram(aes(x = na_prop_pre*100), bins = 12) +
          xlab("Percentage of Missing Values (%)") +
          ylab("Number of Sensors") + theme_g +
          geom_vline(xintercept = 10, size = 2,
                     colour = 'red') +
          geom_label(aes(label = "10% Threshold", x = 16, y = 25),
                    colour = "white", fill = 'black', alpha = 0.4) +
          scale_y_continuous(limits = c(0,35))
grid.arrange(p2, p1, nrow = 1)
```

@@

```{r, fig.width=3, fig.height=3, fig.align='center', fig.cap = "Map showing the algorithm selecting Flinders St Station Underpass and Sandridge Bridge as neighbours to be used to predict the pedestrian counts at Southbank"}
ped_loc <- read_csv("../data/Pedestrian_sensor_locations.csv")
lats <- ped_loc$Latitude
lons <- ped_loc$Longitude
ids <- ped_loc$`Sensor Description`

@@

```{r, echo = FALSE, message = FALSE, warning = F, fig.width = 8, fig.height = 4}
grid.arrange(p1, p2, ncol = 1)
```

@@

## Implementation of improved imputation algorithm

@@

##### Finding neighbouring sensors
Using location data of the sensors provided by the City of Melbourne, a data frame containing the latitude and longitude coordinates of all the sensors which were classified as having a small proportion of missing values (and now have imputed, complete data), which are the candidates to be neighbours used for prediction.  
Using the Great Circle/Haversine distance equation, implemented in the function `gcd.slc()`, an approximation of distance between each of the candidate sensors is calculated in and stored as a vector. Sorting by ascending order (using `sort()`) and taking the two smallest distances with `head(., 2)` returns the two geographically closest candidate neighbours.

@@

```{r, fig.height=4, fig.width=9, fig.cap = "A plot of the in-sample Actual vs Predicted pedestrian counts at Southern Cross Station (sensor location with small proportion of missing values) for the simple GLM model and the improved GLM model. The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points."}
p1 <- ggplot(data = NULL, aes(x = original_df %>% 
             filter(Date_Time < "2017-01-01") %>%
             .[, "Southern Cross Station"],
           y = fitted_base$`Southern Cross Station`)) +
           geom_point(alpha = 0.05) +
           geom_abline(aes(slope = 1, intercept = 0),
                       colour = viridis(3)[1],
                       alpha = 0.7, size = 1.25,
                       linetype = 2) +
           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
                       colour = viridis(3)[3]) +
           theme_minimal() + 
           theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5)) +
           scale_x_continuous(limits = c(0, 4000)) +
           scale_y_continuous(limits = c(0, 4000)) +
           xlab("Actual Pedestrian Count") +
           ylab("Predicted Pedestrian Count") +
           ggtitle("Simple GLM Model: Actual v. Predicted \n at Southern Cross Station")
p2 <- ggplot(data = NULL, aes(x = original_df %>% 
             filter(Date_Time < "2017-01-01") %>%
             .[, "Southern Cross Station"],
           y = fitted_data$`Southern Cross Station`)) +
           geom_point(alpha = 0.05) +
           geom_abline(aes(slope = 1, intercept = 0),
                       colour = viridis(3)[1],
                       alpha = 0.7, size = 1.25,
                       linetype = 2) +
           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
                       colour = viridis(3)[3]) +
           theme_minimal() + 
           theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5)) +
           scale_x_continuous(limits = c(0, 4000)) +
           scale_y_continuous(limits = c(0, 4000)) +
           xlab("Actual Pedestrian Count") +
           ylab("Predicted Pedestrian Count") +
           ggtitle("Improved GLM Model: Actual v. Predicted \n at Southern Cross Station")
grid.arrange(p1, p2, nrow = 1)
```

@@

```{r, fig.height=4, fig.width=9, fig.cap = paste("A plot of the in-sample Actual vs Predicted pedestrian counts at Australia on Collins (sensor location with large proportion of missing values) for the simple GLM model and the improved GLM model. The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points.")}

@@

```{r, fig.height=3.5, fig.width=9, fig.cap = paste("Distribution of in-sample (2014-2016 data) MARE by imputation model")}

@@

```{r, fig.height=4, fig.width=9, fig.cap = "A plot of the out-of-sample (2017) Actual vs Predicted pedestrian counts at Melbourne Central (sensor location with small proportion of missing values) for the simple GLM model and the improved GLM model. The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points."}
p1 <- ggplot(data = NULL, aes(x = ori_ts[, "Melbourne Central"],
           y = test_base$`Melbourne Central`)) +
           geom_point(alpha = 0.05) +
           geom_abline(aes(slope = 1, intercept = 0),
                       colour = viridis(3)[1],
                       alpha = 0.7, size = 1.25,
                       linetype = 2) +
           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
                       colour = viridis(3)[3]) +
           theme_minimal() + 
           theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5)) +
           scale_x_continuous(limits = c(0, 4000)) +
           scale_y_continuous(limits = c(0, 4000)) +
           xlab("Actual Pedestrian Count") +
           ylab("Predicted Pedestrian Count") +
           ggtitle("Simple GLM Model: Actual v. Predicted \n at Melbourne Central")
p2 <- ggplot(data = NULL, aes(x =ori_ts[, "Melbourne Central"],
           y = test_fit$`Melbourne Central`)) +
           geom_point(alpha = 0.05) +
           geom_abline(aes(slope = 1, intercept = 0),
                       colour = viridis(3)[1],
                       alpha = 0.7, size = 1.25,
                       linetype = 2) +
           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
                       colour = viridis(3)[3]) +
           theme_minimal() + 
           theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5)) +
           scale_x_continuous(limits = c(0, 4000)) +
           scale_y_continuous(limits = c(0, 4000)) +
           xlab("Actual Pedestrian Count") +
           ylab("Predicted Pedestrian Count") +
           ggtitle("Improved GLM Model: Actual v. Predicted \n at Melbourne Central")
grid.arrange(p1, p2, nrow = 1)
```

@@

```{r, fig.height=4, fig.width=9, fig.cap = "A plot of the out-of-sample (2017) Actual vs Predicted pedestrian counts at Collins Place (South) (sensor location with large proportion of missing values) for the simple GLM model and the improved GLM model. The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points."}
p1 <- ggplot(data = NULL, aes(x = ori_ts[, "Collins Place (South)"],
           y = test_base$`Collins Place (South)`)) +
           geom_point(alpha = 0.05) +
           geom_abline(aes(slope = 1, intercept = 0),
                       colour = viridis(3)[1],
                       alpha = 0.7, size = 1.25,
                       linetype = 2) +
           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
                       colour = viridis(3)[3]) +
           theme_minimal() + 
           theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5)) +
           scale_x_continuous(limits = c(0, 2500)) +
           scale_y_continuous(limits = c(0, 2500)) +
           xlab("Actual Pedestrian Count") +
           ylab("Predicted Pedestrian Count") +
           ggtitle("Simple GLM Model: Actual v. Predicted \n at Collins Place (South)")
p2 <- ggplot(data = NULL, aes(x =ori_ts[, "Collins Place (South)"],
           y = test_fit$`Collins Place (South)`)) +
           geom_point(alpha = 0.05) +
           geom_abline(aes(slope = 1, intercept = 0),
                       colour = viridis(3)[1],
                       alpha = 0.7, size = 1.25,
                       linetype = 2) +
           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
                       colour = viridis(3)[3]) +
           theme_minimal() + 
           theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5)) +
           scale_x_continuous(limits = c(0, 2500)) +
           scale_y_continuous(limits = c(0, 2500)) +
           xlab("Actual Pedestrian Count") +
           ylab("Predicted Pedestrian Count") +
           ggtitle("Improved GLM Model: Actual v. Predicted \n at Collins Place (South)")
grid.arrange(p1, p2, nrow = 1)
```

@@

```{r, fig.height=3.5, fig.width=9, fig.cap = paste("Distribution of out-of-sample (2017 data) MARE by imputation model, excluding Spencer St-Collins St (North)/(South) and Flinders St-Swanston St (West) sensors")}

@@

```{r, fig.width = 8, fig.height = 3, fig.cap="Average pedestrian counts by time of day and type of day (DayType) at Spencer St-Collins St (South) for training data (dark line) and test data (light line)"}
ggplot() + geom_smooth(aes(x = as.integer(Time),
                           y = `Spencer St-Collins St (South)`),
                       colour = viridis(3)[1],
                       data = dfa2) +
           geom_smooth(aes(x = as.integer(Time),
                           y = `Spencer St-Collins St (South)`),
                       colour = viridis(3)[3],
                       data = ori_ts2) +
           theme_minimal() + labs(x = "Time of Day") +
           facet_wrap(~ DayType)
```

@@

```{r, fig.width = 8, fig.height=2, fig.cap = "Plot of simulated missingness method 1 at Bourke Street Mall (South). 20% of the observations are removed from the traning data for the imputation model at random."}
ggplot() + geom_line(aes(x = dfa2$Date_Time, y = mall_missing_1),
                     size = 0.1, alpha = 1) +
           labs(y = "Pedestrian Count", x = "Time and Date") +
          theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
```

@@

```{r, fig.width = 8, fig.height=2, fig.cap = "Plot of simulated missingness method 2 at Bourke Street Mall (South) with 20% of the training data removed as a sequence. This method of simulating missing data is a better representation of the missing data observed at sensors with a large proportion of missing data compared to randomly removing hourly observations from the training data."}
ggplot() + geom_line(aes(x = dfa2$Date_Time, y = mall_missing_2),
                size = 0.1, alpha = 1) +
           labs(y = "Pedestrian Count", x = "Time and Date") +
          theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
```

@@

```{r, fig.width = 8, fig.height=2, fig.cap = "Plot of simulated missingness method 2 at Bourke Street Mall (South) with 50% of the training data removed as a sequence. A large proportion of the training data is removed to test the robustness of the imputation method used at sensors with extremely large proportions of missing data during the training period defined."}
ggplot() + geom_line(aes(x = dfa2$Date_Time, y = mall_missing_3),
                size = 0.1, alpha = 1) +
           labs(y = "Pedestrian Count", x = "Time and Date") +
          theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
```

@@

# Predictive Model

For the purpose of prediction, the use of the most simplistic model possible to make quick predictions of the expected pedestrian counts at each location, accessible to the public. The model is not expected to be able to predict large deviations from time and date based estimates.

Similar to the simple imputation model used,  a generalised linear model with quasi-poisson errors is estimated using time and date based variables as predictors. The model specification used is:

$$E \left[\text{HourlyCounts} \vert  \text{Sensor}, \text{Month}, \text{DayType}, \text{Time} \right] = \exp (\mu_{\text{Sensor}, \text{Month}, \text{DayType}, \text{Time}}) $$

```{r, echo = FALSE}
library(purrr)
library(dplyr)
library(tidyr)
library(lubridate)
library(roxygen2)
load("../data/model_est.RData")

daytype <- function(date){
  if(!is.Date(date)){stop("Object is not a date.")}
  x_day <- wday(date, label = TRUE, abbr = FALSE)
  x_daytype <- ifelse(x_day == "Sunday", "Sunday",
                ifelse(x_day == "Saturday", "Saturday",
                ifelse(x_day == "Monday", "Monday",
                ifelse(x_day == "Friday", "Friday",
                          "Midweek"))))
  return(x_daytype)
}
#' Predicted pedestrian counts at various sensor locations in Melbourne
#'
#' @param pred_date input date
#' @param t_hour input integer
#' @param is_pub_hol input logical
#' @return data frame of the fitted values at each location in wide format, with time and date
#' @export
#' @examples
#' ped_predict()
#' ped_predict("2017-12-25", 18, TRUE)
#' 
ped_predict <- function(pred_date = today(tzone = "Australia/Melbourne"),
                        t_hour = 12,
                        is_pub_hol = FALSE,
                        long = FALSE) {
  if(!(t_hour %in% seq(0, 23))){
    stop("Hour is not between 0:00 and 23:00. Please give hour value between 0 and 23")}
  pred_date <- ymd(pred_date)
  if(!is.Date(pred_date)){
    stop(paste("Date given to predict cannot be parsed as a date. Use the date format 'yyyy-mm-dd'. Example: 25 January 2018 is '2018-01-25'."))
    }

  date_pred <- data.frame(Time = as.factor(t_hour)) %>%
               mutate(DayType = ifelse(is_pub_hol,
                                       "Holiday",
                                       daytype(pred_date)),
                      Month = lubridate::month(pred_date ,label = TRUE, abbr = F),
                      Date = pred_date,
                      Date_Time = ymd_h(paste(Date, t_hour, sep = " ")))

  preds <- numeric()
  for(i in names(pred_model)){
    preds[i] <- predict(pred_model[[i]], newdata = date_pred, type = "response")
  }
  if(long){
    return(preds)
  }
    else
  preds_df <- data.frame("fit" = preds, "Sensor_ID" = names(preds)) %>%
    spread(Sensor_ID, fit) %>% cbind(date_pred) %>% as_tibble

  return(preds_df)
}

ped_predict_day <- function(pred_date = today(tzone = "Australia/Melbourne"),is_pub_hol = FALSE){
  day_df <- data.frame()
      for(t in 0:23){
        t_preds <- ped_predict(pred_date, t_hour = t, is_pub_hol)
        suppressWarnings(day_df <- bind_rows(day_df, t_preds))
      }
  return(as.tibble(day_df))
}

Using a GLM predicts well compared to estimating with other models such as AR($p$) or other time series models when considering the tradeoff between model complexity and prediction accuracy. Due to multiple seasonalities, AR($p$) models would need to be considered on a hour of the week basis at each sensor location. Prediction with time series models would also require of historical data. For the objective of this predictive model, the additional imputs required by end users to predict pedestrian counts when making out of sample predictions is not desirable.

```{r, fig.width=6, fig.height=4, fig.cap = paste("Plot of daily pedestrian counts series at Melbourne Central by DayType and Month. Month of the year is mapped to colour to help identify annual seasonal patterns, where month has an additive effect on the pedestrian counts. At Melbourne Central, different months can be seen to have different mean counts for a given type of day and hour of the day.")}
ggplot(dfa2 %>%
         mutate(Date = date(Date_Time))) +
  geom_path(aes(x = as.integer(Time), y = `Melbourne Central`,
                group = Date, colour = Month),
            alpha = 0.1) +
  scale_colour_viridis(discrete = TRUE, option = "C") +
  facet_wrap(~ DayType) + theme_minimal() +
  theme(legend.position = "none") +
  labs(x = "Time of Day", y = "Pedestrian Counts \n at Melbourne Central")
```

A plot of the daily pedestrian counts series at Melbourne Central by type of day and month show that time and date variables as deterministic predictors adequately explains most variation in the pedestrian counts. It can be seen that month has an additive effect as the pattern of the hourly counts for a given date are not affected by month. The six different factor levels that `DayType`, the type of day, is seen to be a sufficiently explain six different patterns. The type of day, Midweek, shows evidence that no additional information is gained from identifying the day to be Tuesday, Wednesday or Thursday.

```{r, fig.height=4, fig.width=8, fig.cap=paste("Plot of the estimated coefficients of the GLM for the interaction term $\text{DayType} \times \text{Time}$ for each hour at each sensor. Variables (x-axis) is ordered by the scagnostic measure, Outlying. "), include=FALSE, eval = FALSE}
dfcoeff <- data.frame(terms = factor(), coeff = numeric(), sensor = factor())
for(i in names(pred_model)){
  terms <- names(coef(pred_model[[i]]))
  coeff <- coef(pred_model[[i]])
  sensor <- rep(i, length(terms))
  df_b <- data.frame(terms, coeff, sensor)
  dfcoeff <- bind_rows(dfcoeff, df_b)
}
dfcoeff <- dfcoeff %>% spread(terms, coeff)

library(GGally)
library(scagnostics)
ggparcoord(dfcoeff[, 51:74], scale = "std", 
           order = "Clumpy", alphaLines = 0.3) + theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, size = 6)) +
  labs(x = "Parameter")
```

When estimating the GLM at each sensor location with R, only the necessary elements in the model object which are required in order to retain the functionality of R's `predict.glm()` function are saved to file. Saving the entire GLM object as given by the base R `glm()` function is highly inefficient when working with large datasets. This is due to the model object providing redundancies. In particular, elements which are unnecessarily saved for the purposes of the predictive model: the training data, the predictors in a model matrix format (the matrix of $\mathbf{X}$ after being formatted to be used for regression, such as recoding categorical variables into dummy variables), the response variable as another vector ($\mathbf{Y}$), the fitted values of the model after estimation, the residuals of the estimates, as well as the estimated effects.

The result of removing all the unnecessary redundant data elements in the GLM object is a storage size reduction from `r mall_model_2 %>% object.size() %>% format(units = "auto")` to `r mall_model_2 %>% stripGlmLR() %>% object.size() %>% format(units = "auto")`. This is only `r round((mall_model_2 %>% stripGlmLR() %>% object.size() %>% as.numeric/ mall_model_2 %>% object.size() %>% as.numeric)*100, 2)`% of the size of the original GLM object. With `r length(unique(ped_data$Sensor))` sensors (and models), this represents a reduction in model object size from `r format((mall_model_2 %>% object.size()) * length(unique(ped_data$Sensor)), units = "auto")` to `r format((mall_model_2 %>% stripGlmLR() %>% object.size()) * length(unique(ped_data$Sensor)), units = "auto")`.

The estimated model parameters are saved so that they can be used for on-the-fly predictions by end users without needing to estimate the model themselves. Using the estimated model parameters, a new function is written to facilitate the predictions in an easy to use format.

```{r, echo = TRUE, eval = FALSE}
ped_predict(pred_date = "2017-12-26", t_hour = 13, is_pub_hol = TRUE)
```
This function will return tibble (trimmed down version of `data.frame()` in R) containing the predicted counts at each sensor location for the 13th hour (13:00/1:00 PM) on Boxing Day (26 December) 2017. This tibble can then be easily used for visualisations or analysis. Due to the lack of data on future public holiday dates, the function assumes the date given is not a public holiday unless `is_pub_hol = TRUE` is provided.  
Another function which uses the `ped_predict()`, `ped_predict_day()`, returns a tibble containing the predicted counts at each sensor location for all 24 hours of the date given to predict.

Using these function as a basis, visualisations of pedestrian counts can be easily generated in R:
```{r, echo = TRUE, fig.cap = "Prediction for 21/05/2018 in R using `ped_predict_day()`, then plotting the time series using `ggplot2`. This shows the user friendly nature of the function, where the function returns predictions in a tidy, wide format. No data manipulation is required in order to plot the predicted counts.", fig.height=2}
ped_predict_day("2018-05-21") %>% 
  ggplot() + geom_line(aes(x = Date_Time, y = Southbank))
```

>>>>

---
chapter: 1
knit: "bookdown::render_book"
---
---
title: "Analysing Melbourne CBD pedestrian counts data with a focus on imputation methods"
author: "Gavin Chin"
date: "21 September 2017"
output:
  pdf_document: default
  html_document: default
  word_document: default
fontsize: 12pt
---

@@

dfa$DayType <-       ifelse(dfa$Day == "Sunday", "Sunday",
                     ifelse(dfa$Day == "Saturday", "Saturday",
                    ifelse(dfa$Day == "Monday", "Monday",
                     ifelse(dfa$Day == "Friday", "Friday",
                                  "Midweek"))))
dfa$DayType[dfa$IsHDay == 1] <- "Holiday"
dfa$HDay <- ifelse(dfa$IsHDay == 1, "Holiday", dfa$Day)
original_df <- dfa
dfa <- filter(dfa, Date_Time < "2017-01-01")
ori_2 <- filter(original_df, Date_Time < "2017-01-01")
ori_ts <- filter(original_df, Date_Time > "2016-12-31")
ori_ts2 <- ori_ts
```
# Background

### The Data: City of Melbourne Pedestrian Data

@@

# Literature Review
<!-- #### IMPUTATION OF MISSING CLASSIFIED TRAFFIC DATA DURING WINTER SEASON - _Hyuk-Jae Roh, Satish Sharma, Prasanta K. Sahu_   -->

@@

# Methodology: Imputation of Missing Values

@@

```{r naprop_hist, fig.width=6, fig.height=3, fig.cap="Histograms of the proportions of missing values at each sensor during 2014 - 2016 when calculated before and after correction for false zero counts. The distributions of proportion of missing values at each sensor is significantly different after correcting for false zero counts in the data. This emphasises the necessity to correct for false zero counts as it will affect "}
theme_g <- theme_minimal() +
           theme(plot.title = element_text(hjust = 0.5,
                                           size = 10))
p1 <- ggplot() + geom_histogram(aes(x = na_prop*100), bins = 12) +
          xlab("Percentage of Missing Values (%)") +
          ylab("Number of Sensors") + theme_g +
          geom_vline(xintercept = 10, size = 2,
                     colour = 'red') +
          geom_label(aes(label = "10% Threshold", x = 16, y = 25),
                    colour = "white", fill = 'black', alpha = 0.4) +
          scale_y_continuous(limits = c(0,35))
p2 <- ggplot() + geom_histogram(aes(x = na_prop_pre*100), bins = 12) +
          xlab("Percentage of Missing Values (%)") +
          ylab("Number of Sensors") + theme_g +
          geom_vline(xintercept = 10, size = 2,
                     colour = 'red') +
          geom_label(aes(label = "10% Threshold", x = 16, y = 25),
                    colour = "white", fill = 'black', alpha = 0.4) +
          scale_y_continuous(limits = c(0,35))
grid.arrange(p2, p1, nrow = 1)
```

@@

```{r, fig.width=4, fig.height=4, fig.align='center', fig.cap = "Map showing the algorithm selecting Flinders St Station Underpass and Sandridge Bridge as neighbours to be used to predict the pedestrian counts at Southbank"}
ped_loc <- read_csv("../data/Pedestrian_sensor_locations.csv")
lats <- ped_loc$Latitude
lons <- ped_loc$Longitude
ids <- ped_loc$`Sensor Description`

@@

```{r, echo = FALSE, message = FALSE, warning = F, fig.width = 6, fig.height = 3}
grid.arrange(p1, p2, ncol = 1)
```

@@

# Implementation of improved imputation algorithm

@@

##### Finding neighbouring sensors
Using location data of the sensors provided by the City of Melbourne, a data frame containing the latitude and longitude coordinates of all the sensors which were classified as having a small proportion of missing values (and now have imputed, complete data), which are the candidates to be neighbours used for prediction.  
Using the Great Circle/Haversine distance equation, implemented in the function `gcd.slc()` from , an approximation of distance between each of the candidate sensors is calculated in and stored as a vector. Sorting by ascending order (using `sort()`) and taking the two smallest distances with `head(., 2)` returns the two geographically closest candidate neighbours.

@@

```{r, fig.cap = "A plot of the in-sample Actual vs Predicted pedestrian counts at Southern Cross Station (sensor location with small proportion of missing values) for the simple GLM model and the improved GLM model. The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points."}
p1 <- ggplot(data = NULL, aes(x = original_df %>% 
             filter(Date_Time < "2017-01-01") %>%
             .[, "Southern Cross Station"],
           y = fitted_base$`Southern Cross Station`)) +
           geom_point(alpha = 0.05) +
           geom_abline(aes(slope = 1, intercept = 0),
                       colour = viridis(3)[1],
                       alpha = 0.7, size = 1.25,
                       linetype = 2) +
           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
                       colour = viridis(3)[3]) +
           theme_minimal() + 
           theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5)) +
           scale_x_continuous(limits = c(0, 4000)) +
           scale_y_continuous(limits = c(0, 4000)) +
           xlab("Actual Pedestrian Count") +
           ylab("Predicted Pedestrian Count")
p2 <- ggplot(data = NULL, aes(x = original_df %>% 
             filter(Date_Time < "2017-01-01") %>%
             .[, "Southern Cross Station"],
           y = fitted_data$`Southern Cross Station`)) +
           geom_point(alpha = 0.05) +
           geom_abline(aes(slope = 1, intercept = 0),
                       colour = viridis(3)[1],
                       alpha = 0.7, size = 1.25,
                       linetype = 2) +
           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
                       colour = viridis(3)[3]) +
           theme_minimal() + 
           theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5)) +
           scale_x_continuous(limits = c(0, 4000)) +
           scale_y_continuous(limits = c(0, 4000)) +
           xlab("Actual Pedestrian Count") +
           ylab("Predicted Pedestrian Count")
grid.arrange(p1, p2, nrow = 1)
```

@@

```{r, fig.cap = paste("A plot of the in-sample Actual vs Predicted pedestrian counts at Australia on Collins (sensor location with large proportion of missing values) for the simple GLM model and the improved GLM model. The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points.")}

@@

```{r, fig.height=1, fig.width=6, fig.cap = paste("Distribution of in-sample (2014-2016 data) MARE by imputation model")}

@@

```{r, fig.height=3, fig.width=6, fig.cap = "A plot of the out-of-sample (2017) Actual vs Predicted pedestrian counts at Melbourne Central (sensor location with small proportion of missing values) for the simple GLM model and the improved GLM model. The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points."}
p1 <- ggplot(data = NULL, aes(x = ori_ts[, "Melbourne Central"],
           y = test_base$`Melbourne Central`)) +
           geom_point(alpha = 0.05) +
           geom_abline(aes(slope = 1, intercept = 0),
                       colour = viridis(3)[1],
                       alpha = 0.7, size = 1.25,
                       linetype = 2) +
           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
                       colour = viridis(3)[3]) +
           theme_minimal() + 
           theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5)) +
           scale_x_continuous(limits = c(0, 4000)) +
           scale_y_continuous(limits = c(0, 4000)) +
           xlab("Actual Pedestrian Count") +
           ylab("Predicted Pedestrian Count") +
           ggtitle("Simple GLM Model: Actual v. Predicted \n at Melbourne Central")
p2 <- ggplot(data = NULL, aes(x =ori_ts[, "Melbourne Central"],
           y = test_fit$`Melbourne Central`)) +
           geom_point(alpha = 0.05) +
           geom_abline(aes(slope = 1, intercept = 0),
                       colour = viridis(3)[1],
                       alpha = 0.7, size = 1.25,
                       linetype = 2) +
           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
                       colour = viridis(3)[3]) +
           theme_minimal() + 
           theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5)) +
           scale_x_continuous(limits = c(0, 4000)) +
           scale_y_continuous(limits = c(0, 4000)) +
           xlab("Actual Pedestrian Count") +
           ylab("Predicted Pedestrian Count") +
           ggtitle("Improved GLM Model: Actual v. Predicted \n at Melbourne Central")
grid.arrange(p1, p2, nrow = 1)
```

@@

```{r, fig.height=3, fig.width=6, fig.cap = "A plot of the out-of-sample (2017) Actual vs Predicted pedestrian counts at Collins Place (South) (sensor location with large proportion of missing values) for the simple GLM model and the improved GLM model. The dashed line is $x=y$, representing a reference for perfect fit, while solid line is the linear fit of the points."}
p1 <- ggplot(data = NULL, aes(x = ori_ts[, "Collins Place (South)"],
           y = test_base$`Collins Place (South)`)) +
           geom_point(alpha = 0.05) +
           geom_abline(aes(slope = 1, intercept = 0),
                       colour = viridis(3)[1],
                       alpha = 0.7, size = 1.25,
                       linetype = 2) +
           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
                       colour = viridis(3)[3]) +
           theme_minimal() + 
           theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5)) +
           scale_x_continuous(limits = c(0, 2500)) +
           scale_y_continuous(limits = c(0, 2500)) +
           xlab("Actual Pedestrian Count") +
           ylab("Predicted Pedestrian Count") +
           ggtitle("Simple GLM Model: Actual v. Predicted \n at Collins Place (South)")
p2 <- ggplot(data = NULL, aes(x =ori_ts[, "Collins Place (South)"],
           y = test_fit$`Collins Place (South)`)) +
           geom_point(alpha = 0.05) +
           geom_abline(aes(slope = 1, intercept = 0),
                       colour = viridis(3)[1],
                       alpha = 0.7, size = 1.25,
                       linetype = 2) +
           geom_smooth(method = "lm", size = 1.25, alpha = 0.7, 
                       colour = viridis(3)[3]) +
           theme_minimal() + 
           theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5)) +
           scale_x_continuous(limits = c(0, 2500)) +
           scale_y_continuous(limits = c(0, 2500)) +
           xlab("Actual Pedestrian Count") +
           ylab("Predicted Pedestrian Count") +
           ggtitle("Improved GLM Model: Actual v. Predicted \n at Collins Place (South)")
grid.arrange(p1, p2, nrow = 1)
```

@@

```{r, fig.height=1, fig.width=6, fig.cap = paste("Distribution of out-of-sample (2017 data) MARE by imputation model, excluding Spencer St-Collins St (North)/(South) and Flinders St-Swanston St (West) sensors")}

@@

```{r, fig.width = 6, fig.height = 3, fig.cap="Average pedestrian counts by time of day and type of day (DayType) at Spencer St-Collins St (South) for training data (dark line) and test data (light line)"}
ggplot() + geom_smooth(aes(x = as.integer(Time),
                           y = `Spencer St-Collins St (South)`),
                       colour = viridis(3)[1],
                       data = dfa2) +
           geom_smooth(aes(x = as.integer(Time),
                           y = `Spencer St-Collins St (South)`),
                       colour = viridis(3)[3],
                       data = ori_ts2) +
           theme_minimal() + labs(x = "Time of Day") +
           facet_wrap(~ DayType)
```

@@

```{r, fig.width = 5, fig.height=2, fig.cap = "Plot of simulated missingness method 1 at Bourke Street Mall (South). 20% of the observations are removed from the traning data for the imputation model at random."}
ggplot() + geom_line(aes(x = dfa2$Date_Time, y = mall_missing_1),
                     size = 0.1, alpha = 1) +
           labs(y = "Pedestrian Count", x = "Time and Date") +
          theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
```

@@

```{r, fig.width = 5, fig.height=2, fig.cap = "Plot of simulated missingness method 2 at Bourke Street Mall (South) with 20% of the training data removed as a sequence. This method of simulating missing data is a better representation of the missing data observed at sensors with a large proportion of missing data compared to randomly removing hourly observations from the training data."}
ggplot() + geom_line(aes(x = dfa2$Date_Time, y = mall_missing_2),
                size = 0.1, alpha = 1) +
           labs(y = "Pedestrian Count", x = "Time and Date") +
          theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
```

@@

```{r, fig.width = 5, fig.height=2, fig.cap = "Plot of simulated missingness method 2 at Bourke Street Mall (South) with 50% of the training data removed as a sequence. A large proportion of the training data is removed to test the robustness of the imputation method used at sensors with extremely large proportions of missing data during the training period defined."}
ggplot() + geom_line(aes(x = dfa2$Date_Time, y = mall_missing_3),
                size = 0.1, alpha = 1) +
           labs(y = "Pedestrian Count", x = "Time and Date") +
          theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
```

@@
